---
title: "HW 10"
output: pdf_document
---
# Intro to Data Science - HW 10

##### Copyright 2021, Jeffrey Stanton, Jeffrey Saltz, and Jasmina Tacheva


```{r}
# Enter your name here: Chaithra Kopparam Cheluvaiah
```

### Attribution statement: (choose only one and delete the rest)


```{r}
# 1. I did this homework by myself, with help from the book and the professor.
```

**Association mining** can be applied to many data problems beyond the well-known example of **finding relationships between different products in customer shopping data**. In this homework assignment, we will explore **real data** from the banking sector and look for **patterns associated with the likelihood of responding positively to a direct marketing campaign and signing up for a term deposit with the bank (stored in the variable “y”)**. <br>
You can find out more about the variables in this dataset here: https://archive.ics.uci.edu/ml/datasets/bank+marketing

## Part 1: Explore Data Set

A.	Read the contents of the following URL to a dataframe called **bank** <br>
https://intro-datascience.s3.us-east-2.amazonaws.com/bank-full.csv

**Hint**: Even though this is a .csv file, chances are R won’t be able to read it in correctly using the read_csv() function. If you take a closer look at the contents of the URL file, you may notice each field is separated by a **semicolon** (;) rather than a comma. 

In situations like this, consider using either read.csv or read.table, with two additional parameters. *sep=";"* defines how the data is seperated (the default is a comma), and *header=TRUE* defines that there is a header line in the dataset. 


```{r}
# loading the CSV data to dataframe
bank <- read.csv("https://intro-datascience.s3.us-east-2.amazonaws.com/bank-full.csv", sep=";", header=TRUE)
summary(bank)
```

Make sure there are **41,188** rows and **21** columns in your **bank** df.


```{r}
dim(bank) # checking for number of rows and columns
```

B.	Next, we will focus on some key factor variables from the dataset, and convert a few numeric ones to factor variables. Execute the following command.  Write a comment describing how the conversion for each numeric variable works and what are the variables in the resulting dataframe.


```{r}
bank_new <- data.frame(job=as.factor(bank$job),
                     marital=as.factor(bank$marital),
                     housing_loan=as.factor(bank$housing),
                     young=as.factor(bank$age<median(bank$age)),
                     contacted_more_than_once=as.factor(bank$campaign>1),
                     contacted_before_this_campaign=as.factor(bank$previous<0),
                     success=as.factor(bank$y))
# job: converting string to categorical/factor variable. it represents type of job ('admin.','blue-collar',
# 'entrepreneur', 'housemaid','management','retired','self-employed','services','student','technician','unemployed',
# 'unknown')

# marital: converting string to factor variable. it represents marital status (categorical: 'divorced','married',
# 'single','unknown'; note: 'divorced' means divorced or widowed)

# housing_loan: converting string to factor variable. it represenets whether client has housing loan?
# (categorical: 'no','yes','unknown')

# young: converting quantitative variable to factor variable(TRUE/FALSE) based on a condition whether client age is
# less than median age.

# contacted_more_than_once: converting quantitative variable to factor variable (TRUE/FALSE) based on
# number of contacts performed during the campaign. if campaign has value more than 1, then factor variable
# will be TRUE otherwise FALSE

#  contacted_before_this_campaign:converting quantitative variable to factor variable (TRUE/FALSE) based on
# number of contacts performed before the campaign. if previous has value less than zero, then factor vaiable will be
# TRUE otherwise FALSE

# success: converting string variable to factor variable. it represents whether the client subscribed for term deposit?
# (binary: 'yes','no')
```

C.	Count the number of successful term deposit sign-ups, using the table( ) command on the **success** variable.


```{r}
table(bank_new$success) # creating contingency table
```

D.	Express the results of problem C as percentages by sending the results of the table( ) command into the prop.table( ) command.


```{r}
prop.table(table(bank_new$success)) # table with proportions
```

E.	Using the same techniques, show the percentages for the **marital** and **housing_loan** variables as well.


```{r}
prop.table(table(bank_new$marital))
prop.table(table(bank_new$housing_loan))
```

## Part 2: Coerce the data frame into transactions

F.	Install and library two packages: **arules** and **arulesViz**.


```{r}
# importing the required libraries
library(arules)
library(arulesViz)
```

G.	Coerce the **bank_new** dataframe into a **sparse transactions matrix** called **bankX**.


```{r}
bankX <- as(bank_new, "transactions") # converting dataframe to transactions
summary(bankX)
```

H.	Use the itemFrequency( ) and itemFrequencyPlot( ) commands to explore the contents of **bankX**. What do you see?
```{r}
itemFreq <- itemFrequency(bankX)  # relative frequency of occurrence of all attributes in bank_new data
str(itemFreq)
head(itemFreq)
tail(itemFreq)
```
```{r}
itemFrequencyPlot(bankX, topN=20)
# bar graph that shows the most frequently repeated attribute in the left-most bar
```

I.	This is a fairly large dataset, so we will explore only the first 10 observations in the **bankX** transaction matrix:
```{r}
inspect(bankX[1:10]) 
```

Explain the difference between **bank_new** and **bankX** in a block comment:
```{r}
# bank_new is a data frame with rows and columns
# rows will be observation, columns are characteristics of the data
# content of this matrix will be data-point in the format string, floating-point, integer, boolean etc.,

# bankX is a itemMatrix object in sparse format. Each row is a list of items like {job=blue-collar,
#       marital=married,
#       housing_loan=no,
#       young=FALSE,
#       contacted_more_than_once=FALSE,
#       contacted_before_this_campaign=FALSE,
#       success=no}
# The columns are the individual items.
# also, content of this matrix will be only 0 or 1
```

## Part 3: Use arules to discover patterns

**Support** is the proportion of times that a particular set of items occurs relative to the whole dataset. <br>
**Confidence** is proportion of times that the consequent occurs when the antecedent is present. <br>

J.	Use **apriori** to generate a set of rules with support over 0.005 and confidence over 0.3, and trying to predict who successfully signed up for a term deposit. <br>
**Hint:** You need to define the **right-hand side rule (rhs)**.
```{r}
# running the algorithm to find association rules in the transaction data
ruleset <- apriori(bankX,parameter=list(supp=0.005, conf=0.3),
                  control=list(verbose=F),
                  appearance=list(default="lhs",rhs=("success=yes")))
summary(ruleset)
```

K.	Use inspect() to review of the **ruleset**.
```{r}
inspect(ruleset)
```

L.	Use the output of inspect( ) or inspectDT( ) and describe **any 2 rules** the algorithm found.
```{r}
# I picked the rules that has higest lift and having highest confidence among the rules that are returned in
# previous step

# Rule 8: {job=student, marital=single, young=TRUE, contacted_before_this_campaign=FALSE} => {success=yes}
# It the highest lift (2.87) and 32.34% confidence
# whenever {job=student, marital=single, young=TRUE, contacted_before_this_campaign=FALSE} are together
# then "{success=yes}" also appears
# Interpretation: client who is a student, single, young (age<38, median age) and was not contacted before the
# campaign is more likely to subscribe to term deposit.

# Rule 5: {job=student,marital=single,young=TRUE}  => {success=yes}
# It the highest lift (2.87) and 32.34% confidence
# whenever {job=student,marital=single,young=TRUE} are together then "{success=yes}" also appears
# Interpretation: client who is a student, single and young (age<38, median age) is more likely to subscribe
# to term deposit.

# marketing campaigns should be more targeted towards the above individuals


```
