# Intro to Data Science - HW 10

##### Copyright 2021, Jeffrey Stanton, Jeffrey Saltz, and Jasmina Tacheva


```{r}
# Enter your name here: Chaithra Kopparam Cheluvaiah
```

### Attribution statement: (choose only one and delete the rest)


```{r}
# 1. I did this homework by myself, with help from the book and the professor.
# 2. I did this homework with help from the book and the professor and these Internet sources:
# 3. I did this homework with help from <Name of another student> but did not cut and paste any code.
```

**Association mining** can be applied to many data problems beyond the well-known example of **finding relationships between different products in customer shopping data**. In this homework assignment, we will explore **real data** from the banking sector and look for **patterns associated with the likelihood of responding positively to a direct marketing campaign and signing up for a term deposit with the bank (stored in the variable “y”)**. <br>
You can find out more about the variables in this dataset here: https://archive.ics.uci.edu/ml/datasets/bank+marketing

## Part 1: Explore Data Set

A.	Read the contents of the following URL to a dataframe called **bank** <br>
https://intro-datascience.s3.us-east-2.amazonaws.com/bank-full.csv

**Hint**: Even though this is a .csv file, chances are R won’t be able to read it in correctly using the read_csv() function. If you take a closer look at the contents of the URL file, you may notice each field is separated by a **semicolon** (;) rather than a comma. 

In situations like this, consider using either read.csv or read.table, with two additional parameters. *sep=";"* defines how the data is seperated (the default is a comma), and *header=TRUE* defines that there is a header line in the dataset. 


```{r}

```

Make sure there are **41,188** rows and **21** columns in your **bank** df.


```{r}

```

B.	Next, we will focus on some key factor variables from the dataset, and convert a few numeric ones to factor variables. Execute the following command.  Write a comment describing how the conversion for each numeric variable works and what are the variables in the resulting dataframe.


```{r}
bank_new <- data.frame(job=as.factor(bank$job),
                     marital=as.factor(bank$marital),
                     housing_loan=as.factor(bank$housing),
                     young=as.factor(bank$age<median(bank$age)),
                     contacted_more_than_once=as.factor(bank$campaign>1),
                     contacted_before_this_campaign=as.factor(bank$previous<0),
                     success=as.factor(bank$y))
```

C.	Count the number of successful term deposit sign-ups, using the table( ) command on the **success** variable.


```{r}

```

D.	Express the results of problem C as percentages by sending the results of the table( ) command into the prop.table( ) command.


```{r}

```

E.	Using the same techniques, show the percentages for the **marital** and **housing_loan** variables as well.


```{r}

```

## Part 2: Coerce the data frame into transactions

F.	Install and library two packages: **arules** and **arulesViz**.


```{r}

```

G.	Coerce the **bank_new** dataframe into a **sparse transactions matrix** called **bankX**.


```{r}

```

H.	Use the itemFrequency( ) and itemFrequencyPlot( ) commands to explore the contents of **bankX**. What do you see?
```{r}

```

I.	This is a fairly large dataset, so we will explore only the first 10 observations in the **bankX** transaction matrix:
```{r}
inspect(bankX[1:10]) 
```

Explain the difference between **bank_new** and **bankX** in a block comment:
```{r}

```

## Part 3: Use arules to discover patterns

**Support** is the proportion of times that a particular set of items occurs relative to the whole dataset. <br>
**Confidence** is proportion of times that the consequent occurs when the antecedent is present. <br>

J.	Use **apriori** to generate a set of rules with support over 0.005 and confidence over 0.3, and trying to predict who successfully signed up for a term deposit. <br>
**Hint:** You need to define the **right-hand side rule (rhs)**.
```{r}

```

K.	Use inspect() to review of the **ruleset**.
```{r}

```

L.	Use the output of inspect( ) or inspectDT( ) and describe **any 2 rules** the algorithm found.
```{r}

```
