# Intro to Data Science - HW 5

##### Copyright 2021, Jeffrey Stanton, Jeffrey Saltz, and Jasmina Tacheva


```{r}
# Enter your name here: 
```

### Attribution statement: (choose only one and delete the rest)


```{r}
# 1. I did this homework by myself, with help from the book and the professor.
# 2. I did this homework with help from the book and the professor and these Internet sources:
# 3. I did this homework with help from <Name of another student> but did not cut and paste any code.
```

Reminders of things to practice from previous weeks: <br>
Descriptive statistics: 	mean( ) max( ) min( )<br>
Coerce to numeric:		as.numeric( )<br>

## Part 1: Use the Starter Code 

Below, I have provided a starter file to help you. <br>

Each of these lines of code **must be commented** (the comment must that explains what is going on, so that I know you understand the code and results).


```{r}
library(RCurl)
library(jsonlite)
dataset <- getURL("https://intro-datascience.s3.us-east-2.amazonaws.com/role.json")
readlines <- jsonlite::fromJSON(dataset)
df <- readlines$objects$person
```

A.	Explore the **df** dataframe (e.g., using head() or whatever you think is best).


```{r}

```

B.	Explain the dataset <br>
o	What is the dataset about? <br>
o	How many rows are there and what does a row represent?<br>
o	How many columns and what does each column represent?<br>



```{r}

```

## Part 2: Investigate the resulting dataframe 

C.	How many senators are women? 


```{r}

```

D.	How many senators have a YouTube account?  


```{r}

```

E.	How many women senators have a YouTube account?


```{r}

```

F.	Create a new dataframe called **youtubeWomen** that only includes women senators who have a YouTube account.   


```{r}

```

G.	What does running this line of code do? Explain in a comment: 


```{r}
youtubeWomen$year <- substr(youtubeWomen$birthday,1,4)
```

H.	Use this new variable to calculate the mean **birthyear** in **youtubeWomen**. <br>
**Hint:** You may need to convert it to numeric first.


```{r}

```

I.	Make a histogram of the **birthyears** of senators in **youtubeWomen**. Add a comment describing the shape of the distribution.


```{r}

```
