---
title: "Lab 9"
output: pdf_document
---

```{r}
# 1. I did this homework by myself, with help from the book and the professor.

# installing packages
install.packages("kernlab")
install.packages("caret")
```
```{r}
library(kernlab)
library(caret)
data("GermanCredit")
subCredit <- GermanCredit[,1:10]
str(subCredit)

```
1.	Examine the data structure that str() reveals. Also use the help() function to learn more about the GermanCredit data set. Summarize what you see in a comment.
```{r}
help(GermanCredit)
str(subCredit)
```

2.	Use the createDataPartition() function to generate a list of cases to include in the training data. This function is conveniently provided by caret and allows one to directly control the number of training cases. It also ensures that the training cases are balanced with respect to the outcome variable. Try this: trainList <- createDataPartition(y=subCredit$Class,p=.40,list=FALSE)
```{r}
trainList <- createDataPartition(y=subCredit$Class,p=.40,list=FALSE)
```

3.	Examine the contents of trainList to make sure that it is a list of case numbers. With p=0.40, it should have 400 case numbers in it.
```{r}

```

4.	What is trainList? What do the elements in trainList represent? Which attribute is balanced in the trainList dataset?
```{r}

```

5.	Use trainList and the square brackets notation to create a training data set called “trainSet” from the subCredit data frame. Look at the structure of trainSet to make sure it has all of the same variables as subCredit. The trainSet structure should be a data frame with 400 rows and 10 columns.
```{r}
trainSet <- subCredit[trainList,]
```

6.	Use trainList and the square brackets notation to create a testing data set called “testSet” from the subCredit data frame. The testSet structure should be a data frame with 600 rows and 10 columns and should be a completely different set of cases than trainSet.
```{r}
testSet <- subCredit[-trainList,]
```
7.	Create and interpret boxplots of all the predictor variables in relation to the outcome variable (Class).
```{r}

```
8.	Train a support vector machine with the ksvm() function from the kernlab package. Make sure that you have installed and libraried the kernlab package. Have the cost be 5, and have ksvm do 3 cross validations (hint: try prob.model = TRUE)
```{r}

```

9.	Examine the ksvm output object. In particular, look at the cross-validation error for an initial indication of model quality. Add a comment that gives your opinion on whether this is a good model.
```{r}

```

10.	Predict the training cases using the predict command
```{r}

```
11.	Examine the predicted out object with str( ). Then, calculate a confusion matrix using the table function.
```{r}

```
12.	Interpret the confusion matrix and in particular calculate the overall accuracy of the model. The diag( ) command can be applied to the results of the table command you ran in the previous step. You can also use sum( ) to get the total of all four cells.
```{r}

```

13.	Check you calculation with confusionMatrix() function in the caret package.
```{r}

```
